---
title: "Activity Monitor"
author: "Nan Zhong"
date: "Sunday, October 19, 2014"
output: html_document
---
Daily Foot steps Analysis
==========================================================
Load the raw data, the necessary libraries for processing data, and convert data into data table format.

```{r,results='hide'}
setwd("C:/Users/nzhong/Documents/rstudio/CourseraDataScienceTrack/repreducible Research/")
data<-read.csv("activity.csv")
library(data.table)
library(ggplot2)
options(scipen = 999)
no_na_data<-as.data.table(data[complete.cases(data),])
```

## Part I  Ignore Missing Records##

### 1.1 Analysis of number of steps taken per day
```{r,results='hide'}
daily_steps<-no_na_data[,sum(steps),by=list(date)]$V1
hist(daily_steps,breaks=20,xlab="number of steps per day",main="Daily Steps Histogram(ignore missing value)",ylim=c(0,15))
```
  
the mean and median total number of steps taken per day are **`r mean(daily_steps)`** and **`r median(daily_steps)`** separately

### 1.2 Analysis of average daily activity pattern
```{r}
daily_patterns<-no_na_data[,mean(steps),by=list(interval)]
setnames(daily_patterns,c("interval","AvgSteps"))
max_steps_interval<-daily_patterns[order(-AvgSteps)][1,]
plot(daily_patterns,xlab="daily time stamps",ylab="average number of steps",main="Daily Step Patterns")
```
  
the maximum number of steps, on average across all the days in the dataset, is `r max_steps_interval$AvgSteps`    
and it occurred on the interval `r max_steps_interval$interval`


# Part II Imputing missing values

Calculate the number of missing records
```{r}
num_missing<-dim(data)[1]-dim(no_na_data)[1]
```
There are altogether **`r num_missing`** missing records out of **`r dim(data)[1]`** total records

All the missing step values are replaced by the mean number of steps on average across all the days in the dataset with the same interval. 
```{r,results='hide',include=FALSE}
temp<-data[is.na(data$steps),]
fill<-as.data.table(merge(temp,daily_patterns,by="interval"))
fill[,steps:=NULL]
setnames(fill,"AvgSteps","steps")
new_data<-rbind(fill,no_na_data)
``` 

### 2.1 Draw the histogram of number of steps per day again with this new dataset, we get
```{r,echo=FALSE}
daily_steps_new<-new_data[,sum(steps),by=list(date)]$V1
hist(daily_steps_new,breaks=20,xlab="number of steps per day",main="Daily Steps Histogram(missing value filled)")
```
  
the mean and median total number of steps taken per day are **`r mean(daily_steps_new)`** and **`r median(daily_steps_new)`** separately  
It's almost the same as the ignoring NA case.  
According to the plot, the variations of the frequency is increased after imputing missing value.

### 2.2 Daily step patterns difference between weekdays and weekends
```{r,results='hide'}
new_data$DayOfWeek<-ifelse(!weekdays(as.Date(new_data$date)) %in% c("Saturday", "Sunday"),
                           "weekday","weekend")
daily_patterns_new<-new_data[,mean(steps),by=list(interval,DayOfWeek)]
qplot(interval,V1,data=daily_patterns_new,facets=DayOfWeek ~.,xlab="daily time stamps",ylab="average number of steps",main="Daily Step Patterns")
```